\vspace*{-8mm}\textbf{\textsc{Novelty Detection for Semantic Place Categorization}}

### 2011-06-23 thesis/novelty-intro.tex
Notes on how to train a threshold.

\section{Novelty Detection by Thresholding}
The objective of a novelty detection system is to classify a given sample $x$ as
either \emph{known}: $x$ being generated by a known class to the system, or \emph{novel}:
$x$ generated by a class unknown for the system.
Based on the ground\hyp{}truth of a sample four cases are possible:

@@

\begin{description}
\item[uniform model] - by assuming the factor to be uniform, it can be considered as not existent on
the graphical model and so special methods can be devised to calculate novelty on several variables
by analysing the messages that are propagated in and out of each variable when performing inference.

\item[independent model] - it can very easily be trained from unlabelled data. 
\end{description}

\subsection{Threshold}
After defining a model for the conditional and unconditional distributions, a threshold can be applied
over the ratio of those.

For that

>>>>

\section{Novelty Detection by Thresholding}
\label{sec:threshold}
The objective of a novelty detection system is to classify a given sample $x$ as
either \emph{known}: $x$ being generated by a known class to the system, or \emph{novel}:
$x$ generated by a class unknown for the system.
Based on the ground\hyp{}truth of a sample four cases are possible:

@@

\begin{description}
\item[uniform model] - by assuming the factor to be uniform, it can be considered as not existent on
the graphical model.

\item[independent model] - due to the use of single connected factors, it can easily be trained from
                           unlabelled data. This helps the system to account for an existent bias on
                           each of the variables without allowing the system to overfit to the unlabelled
                           data.
\end{description}

\subsection{Threshold}
After defining a model for the conditional and unconditional distributions, a threshold can be applied
over the ratio of those.
Under the assumption of a constant $P(novel)$, defining a threshold over it leads to an optimal novelty
detector, as showed on \autoref{sec:threshold}.

If the training data does not contains any bias, by picking a threshold on the training data that achieves
a certain true-positive and false-positive rate, it is expected that on the real distribution such a system
will achieve the same performance.

Note however that the performance of the system is dependent on how well the models approximate the conditional
and unconditional distributions and so are dependent on how the graph structure is able to model the real distribution
and how well the training data allows the factors to be approximated.
